---
title: "Week 3 Assignment & Notes"
format: 
    html:
        embed-resources: true
execute:
    echo: true
---

# Week 3 Assignment

## Question 1

```{r}
library(readxl)

myData <- read_excel("../data/Ch11_Q4_Data_File.xlsx")
View(myData)
```

Using the original values, compute the **Euclidean** distance for all possible pairs of the first three observations.

```{r}
euclids <- dist(myData, method = "euclidean")
```

```{r}
euclid_matrix <- as.matrix(euclids)

euclid_matrix[1, 2]
euclid_matrix[1, 3]
euclid_matrix[2, 3]
```

Based on the entire data set, find the minimum and maximum values for the three variables. Use the min-max transformation to normalize the values, and then compute the **Euclidean** distance for all possible pairs of the first three observations.

```{r}
min_values <- apply(myData, 2, min)
max_values <- apply(myData, 2, max)

print("Minimum values:")
print(min_values)
print("Maximum values:")
print(max_values)
```

```{r}
normalize <- function(x) {
    return((x - min(x)) / (max(x) - min(x)))
}
```

```{r}
myData_normalized <- as.data.frame(apply(myData, 2, normalize))
```

```{r}
euclids_norm <- dist(myData_normalized[1:3,], method = "euclidean")
euclid_matrix_norm <- as.matrix(euclids_norm)

euclid_matrix_norm[1,2]
euclid_matrix_norm[1,3]
euclid_matrix_norm[2,3]
```

Using the original values, compute the **Manhattan** distance for all possible pairs of the first three observations.

```{r}
manhattans <- dist(myData, method = "manhattan")
```

```{r}
manhattan_matrix <- as.matrix(manhattans)

manhattan_matrix[1, 2]
manhattan_matrix[1, 3]
manhattan_matrix[2, 3]
```

Using the min-max normalized values, compute the **Manhattan** distance for all possible pairs of the first three observations.

```{r}
manhattans_norm <- dist(myData_normalized[1:3,], method = "manhattan")
manhattan_matrix_norm <- as.matrix(manhattans_norm)

manhattan_matrix_norm[1,2]
manhattan_matrix_norm[1,3]
manhattan_matrix_norm[2,3]
```

## Question 2

```{r}
myData <- read_excel("../data/Ch11_Q10_Data_File.xlsx")
View(myData)
```

Without transforming the values, compute the **Euclidean** distance for all possible pairs of the first three employees.

```{r}
euclids <- dist(myData, method = "euclidean")
```

```{r}
euclid_matrix <- as.matrix(euclids)

euclid_matrix[1, 2]
euclid_matrix[1, 3]
euclid_matrix[2, 3]
```

Compute the z-score standardized salaries and GPAs for the first three employees.

$$
z = \frac{x-\mu}{\sigma}
$$

```{r}
z <- function(val, mean, sd) {
    return ((val - mean) / sd)
}
```

```{r}
mean_salary <- mean(myData$Salary)
sd_salary <- sd(myData$Salary)

myData$ZSalary <- z(myData$Salary, mean_salary, sd_salary)
```

```{r}
mean_gpa <- mean(myData$GPA)
sd_gpa <- sd(myData$GPA)

myData$ZGPA <- z(myData$GPA, mean_gpa, sd_gpa)
```

Based on the z-score standardized salaries and GPAs, compute the **Euclidean** distance for all possible pairs of the first three employees.

```{r}
euclids <- dist(myData[, c("ZSalary", "ZGPA")], method = "euclidean")
```

```{r}
euclid_matrix <- as.matrix(euclids)

euclid_matrix[1, 2]
euclid_matrix[1, 3]
euclid_matrix[2, 3]
```

Based on the z-score standardized salaries and GPAs, compute the **Manhattan** distance for all possible pairs of the first three employees.

```{r}
manhattans <- dist(myData, method = "manhattan")
```

```{r}
manhattan_matrix <- as.matrix(manhattans)

manhattan_matrix[1, 2]
manhattan_matrix[1, 3]
manhattan_matrix[2, 3]
```

Based on the z-score standardized Manhattan distance values, identify the pair of the first three employees that are most similar.

*1 and 3 are the most similar because they have the lowest manhattan distance.*

## Question 3

```{r}
myData <- read_excel("../data/Ch11_Q17_Data_File.xlsx")
View(myData)
```

Using the first five sales transactions, compute the matching coefficients for all pairwise observations for the three binary variables.

$$
\text{Matching Coefficient} = \frac{\text{Number of Variables with Matching Outcomes}}{\text{Total Number of Variables}}
$$

```{r}
matching_coef <- function(row1, row2) {
    matches <- sum(row1 == row2)
    total <- length(row1)
    return(matches / total)
}
```

```{r}
transactions <- myData[1:5, c("Hamburgers", "Fries", "Soda")]
```

```{r}
matching_coef(transactions[1,], transactions[2,])
matching_coef(transactions[1,], transactions[3,])
matching_coef(transactions[1,], transactions[4,])
matching_coef(transactions[1,], transactions[5,])

matching_coef(transactions[2,], transactions[3,])
matching_coef(transactions[2,], transactions[4,])
matching_coef(transactions[2,], transactions[5,])

matching_coef(transactions[3,], transactions[4,])
matching_coef(transactions[3,], transactions[5,])

matching_coef(transactions[4,], transactions[5,])
```

Using the first five sales transactions, compute Jaccard’s coefficients for all pairwise observations for the three binary variables.

```{r}
jaccards <- dist(transactions, method = "binary")
jaccards
```

Based on the matching coefficients, which of the following transactions is the most similar pair?

*for jaccards, we want a smaller value. for matching coefficients, we want a larger value*

*Observations 1 and 5 are most similar;* `matching_coef = 1`

Based on the matching coefficients, which of the following transactions is the least similar pair?

*Observations 2 and 4 are least similar;* `matching_coef = 0`

## Question 4

```{r}
# function to do this all for me
report <- function(df) {
    TP <- sum(df$`Actual Class` == 1 & df$Predicted_Class == 1)
    TN <- sum(df$`Actual Class` == 0 & df$Predicted_Class == 0)
    FP <- sum(df$`Actual Class` == 0 & df$Predicted_Class == 1)
    FN <- sum(df$`Actual Class` == 1 & df$Predicted_Class == 0)

    accuracy <- (TP + TN) / (TP + TN + FP + FN)
    print(paste("Accuracy:", accuracy))

    misclassification_rate <- 1 - accuracy
    print(paste("Misclassification Rate: ", misclassification_rate))

    sensitivity <- (TP) / (TN + FN)
    print(paste("Sensitivity: ", sensitivity))

    precision <- (TP) / (TP + FP)
    print(paste("Precision: ", precision))

    specificity <- (TN) / (TN + FP)
    print(paste("Specificity: ", specificity))

    
}
```

```{r}
myData <- read_excel("../data/Ch11_Q23_Data_File.xlsx")
View(myData)
```

Compute the misclassification rate, accuracy rate, sensitivity, precision, and specificity using the cutoff value of 0.5.

```{r}
# create predicted class based on cutoff
myData$Predicted_Class <- ifelse(myData$`Class 1 Probability` >= 0.5, 1, 0)
```

```{r}
report(myData)
```

Compute the misclassification rate, accuracy rate, sensitivity, precision, and specificity using the cutoff value of 0.25.

```{r}
# create predicted class based on cutoff
myData$Predicted_Class <- ifelse(myData$`Class 1 Probability` >= 0.25, 1, 0)
```

```{r}
report(myData)
```

Compute the misclassification rate, accuracy rate, sensitivity, precision, and specificity using the cutoff value of 0.75.

```{r}
# create predicted class based on cutoff
myData$Predicted_Class <- ifelse(myData$`Class 1 Probability` >= 0.75, 1, 0)
```

```{r}
report(myData)
```

## Question 5

```{r}
myData <- read_excel("../data/Ch11_Q34_Data_File.xlsx")
View(myData)
```

Compute the ME, RMSE, MAD, MPE, and MAPE for the two predictive models.

$$
\text{Mean Error} = ME = \frac{1}{n} \sum^n_{i=1} (y_i - \hat{y}_i)
$$
\text{Mean Absolute Deviation} = MAD = \frac{1}{n} \sum^n_{i=1} |y_i - \hat{y}_i|
$$ 
\text{Root Mean Square Error} = RMSE = \sqrt{\frac{1}{n} \sum^n_{i=1}(y_i - \hat{y}_i)^2}
$$
$$ 
\text{Mean Percentage Error} = MPE = \frac{1}{n} \sum^n_{i=1} \frac{y_i-\hat{y}_i}{y_i} \times 100
$$
$$ 
\text{Mean Absolute Percentage Error} = MAPE = \frac{1}{n} \sum^n_{i=1} |\frac{y_i-\hat{y}_i}{y_i}| \times 100
$$

```{r}
report <- function(actual, predicted) {
    n <- length(actual)
    
    mean_error <- (sum(actual - predicted) / n)
    print(paste("ME: ", mean_error))
    
    mean_absolute_deviation <- (sum(abs(actual - predicted)) / n)
    print(paste("MAD: ", mean_absolute_deviation))
    
    root_mean_square_error <- sqrt(sum((actual - predicted)**2) / n)
    print(paste("RMSE: ", root_mean_square_error))
    
    mean_percentage_error <- (sum(( (actual - predicted) / actual)*100 )) / n
    print(paste("MPE: ", mean_percentage_error))
    
    mean_absolute_percentage_error <- (sum( abs( (actual - predicted) / actual) * 100)) / n
    print(paste("MAPE: ", mean_absolute_percentage_error))
}
```

```{r}
print("Model 1 \n")
report(myData$`Actual Price`, myData$`Predicted Price 1`)
```

```{r}
print("Model 2 \n")
report(myData$`Actual Price`, myData$`Predicted Price 2`)
```

Compare the predictive models to a base model where every house is predicted to be sold at the average price of all the houses in the training data set, which is \$260,500. Compute RMSE for the base model.

```{r}
myData$Average = 260500
```

```{r}
report(myData$`Actual Price`, myData$Average)
```